---
title: "aggregated_figures"
author: "Juliano Palacios Abrantes"
date: "`r Sys.Date()`"
output: html_document
---

This script is used to produce the data and figures for the aggregated results of the scenarios paper. It includes the relative changes in biomass for different scenarios and time periods, as well as comparisons between protected and non-protected areas.


```{r setup, include = F}

# Load functions
source("/Users/jepa88/GitHub/etpmc_cc/functions/mpa_names_fx.R")
source("/Users/jepa88/GitHub/etpmc_cc/functions/scenario_names_fx.R")


library(MyFunctions)

my_lib(c(
  # For grid estimation and species selection
  "tidyverse","sf",
  # For collaborative ease
  "here",
  #For standarize data
  "scales",
  "zoo",
  # For pretty plots
  "wesanderson",
  "tidytext",
  "ggh4x" # faceting edditing
))

# For grid e
# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

select <- dplyr::select


# Change paths to your computer

data_path <- "~/Library/CloudStorage/GoogleDrive-jepa88@gmail.com/My Drive/lenfest_mpa_project/data/"


results_path <- "~/Library/CloudStorage/GoogleDrive-jepa88@gmail.com/My Drive/lenfest_mpa_project/data/03 results/scenarios_paper/"
```

# Aggregated results

## Data load

```{r load_data}

# Load pre created data

scenarios_data_abs <- read_csv(
  paste0(data_path,"04 raw dbem results/processed/all/","all_years_data_abd_esm.csv")
)

# Load sau data 
sau_sf <- my_sf("SAU") %>% 
  filter(name %in% c("Costa Rica (Pacific)","Colombia (Pacific)","Panama (Pacific)","Galapagos Isl. (Ecuador)", "Ecuador"))

# Load CMAR shapefile
cmar_sf <- read_sf(paste0(data_path,"01 spatial/cmar_capas/cmar_shp/all_cmar_mpas.shp")) %>% 
  slice(-4,-5,-9)

```

## Data prep

```{r}
sq_df <- scenarios_data_abs %>% 
  filter(scen == "sq") %>% 
  filter(year < 2014) %>% 
  group_by(region,esm,ssp) %>%
  summarise(mean_time = mean(value_region, na.rm = T)) %>% 
  group_by(region,ssp) %>%
  summarise(mean_esm = mean(mean_time, na.rm = T))


data_df <- scenarios_data_abs %>% 
  mutate(
    time_frame = case_when(
      # year > 1994 & year < 2014 ~ "historic",
      year < 2015 ~ "historic",
      year > 2021 & year < 2040 ~ "early_2030",
      year > 2041 & year < 2060 ~ "mid_2050",
      TRUE ~ NA_character_
    )
  ) %>% 
  filter(!is.na(time_frame)) %>% 
  group_by(region,scen,esm,ssp,time_frame) %>%
  summarise(mean_time = mean(value_region, na.rm = T)) %>% 
  group_by(region,scen,ssp,time_frame) %>% 
  pivot_wider(names_from = time_frame, values_from = mean_time) %>% 
  # left_join(sq_df) %>% 
  mutate(
    dif_2030 = my_chng(historic,early_2030),
    dif_2050 = my_chng(historic,mid_2050),
    # dif_2040sq = my_chng(mean_esm,early_2050),
  ) %>% 
  scenario_names(leng = "eng") %>% 
  # filter(!is.na(dif_2040)) %>% 
  # filter(region != "High seas") %>% 
  summarise(
    across(
      c(dif_2030, dif_2050),
      list(mean_esm = ~mean(.x, na.rm = TRUE),
           sd_esm = ~sd(.x, na.rm = TRUE)),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  ) %>% 
  pivot_longer(cols = c("dif_2030_mean_esm": "dif_2050_sd_esm"),
               names_to = "period",
               values_to = "value") %>% 
  mutate(
    stat = ifelse(str_detect(period, "mean"), "mean_esm", "sd_esm"),
    period = ifelse(str_detect(period, "2030"), "2030", "2050"),
    ssp = ifelse(ssp == "ssp126", "SSP1-2.6","SSP5-8.5")
  )


```

## Fig 3A. Relative to historic period (period)

```{r}

sq_data <- data_df %>% 
  filter(scen == "sq") %>% 
  scenario_names(leng = "eng") %>% 
  spread(stat,value) %>% 
  mutate(
    z_score = mean_esm / sd_esm,
    confidence_flag = case_when(
      abs(z_score) > 2 ~ "high",
      abs(z_score) > 1 ~ "moderate",
      TRUE ~ "low"
    )
  )

# Confidence
sau_points_sf <-
  sau_sf %>%
  rename(region = name) %>% 
  left_join(sq_data) %>%
  select(region:scenario,confidence_flag) %>%
  st_centroid()

# Confidence
cmar_points_sf <-
  cmar_sf %>%
  left_join(sq_data) %>%
  select(region:scenario,confidence_flag) %>%
  st_centroid()


map <-
  ggplot() +
  geom_sf(data = sau_sf %>% select(region = name) %>%  left_join(sq_data),
          aes(fill = mean_esm),
          color = "grey"
  ) +
  geom_sf(data = cmar_sf %>%  left_join(sq_data),
          aes(fill = mean_esm),
          color = "black"
  ) +
  geom_sf(data = sau_points_sf,
          aes(shape = confidence_flag)
  ) +
  geom_sf(data = cmar_points_sf,
          aes(shape = confidence_flag)
  ) +
  facet_grid(ssp~period) +
  # For SQ scenario change
  scale_fill_gradient2("Change in biomass\nrelative to baseline (%)",
                       limits = c(-15,1),
                       breaks = seq(-15,1,5)
  ) +
  my_land_map() +
  my_ggtheme_m("Reg",
               facet_tl_s = 10,
               ax_tl_s = 8,
               ax_tx_s = 8,
               leg_tl_s = 8,
               leg_tx_s = 8,
               leg_pos = "right" 
  ) +
  # theme(
  # legend.key.width = unit(3, "line"),
  # title = element_text(size = 4)
  # )  +
  # guides(fill = guide_colorbar(title.position = "top", title.hjust = 0.5)
  # ) +
  coord_sf(
    xlim = c(-95,-77),
    ylim = c(-5,9)
  )


ggsave(
  paste0(results_path,"figures/relative_hist/all/sq_relative_map.png"),
  # paste0(results_path,"figures/relative_scen/all/ssp126_sq_relative_map.png"),
  map,
  height = 7,
  width = 8
)


```

## Fig 3B. Relative to historic period (Year)

```{r}

hist_df <- scenarios_data_abs %>%
  filter(scen == "sq",
         # region != "High seas"
  ) %>%
  filter(year < 2015) %>%
  group_by(esm,ssp,scen) %>%
  summarise(mean_time = mean(value_region, na.rm = T),
            .groups = "drop" 
  )


esm_mean <- scenarios_data_abs %>% 
  mutate(year = as.numeric(year)) %>%
  filter(
    # region != "High seas",
    scen == "sq") %>%
  group_by(year,esm,ssp,scen) %>%
  summarise(future_time = mean(value_region, na.rm = T),
            .groups = "drop" 
  ) %>%
  left_join(hist_df,
            by = c("esm", "ssp", "scen")
  ) %>% 
  mutate(
    per_change = my_chng(mean_time,future_time)
  )

scenarios_data_abs %>% 
  mutate(year = as.numeric(year)) %>%
  filter(
    # region != "High seas",
    scen == "sq") %>%
  group_by(year,esm,ssp,scen) %>%
  summarise(future_time = mean(value_region, na.rm = T),
            .groups = "drop" 
  ) %>%
  left_join(hist_df,
            by = c("esm", "ssp", "scen")
  ) %>% 
  mutate(
    per_change = my_chng(mean_time,future_time)
  ) %>% 
  group_by(year,ssp,scen) %>%
  mutate(esm_mean = mean(per_change, na.rm = T)) %>% 
  group_by(scen,ssp,esm) %>%
  mutate(
    rolling_mean_sem = rollmean(esm_mean, k = 5, fill = NA, align = "right")
  ) %>%
  # View()
  filter(year > 2015) %>% 
  ggplot() +
  geom_line(
    aes(
      x = year,
      y = per_change,
      color = ssp,
      linetype = esm
    ),
    alpha = 0.5
  ) +
  geom_line(
    aes(
      x = year,
      y = rolling_mean_sem,
      color = ssp
    )
  ) +
  geom_vline(xintercept = 2060, linetype = "dashed", color = "grey")+
  scale_linetype("Earth system model") +
  scale_y_continuous("Change in biomass relative to baseline (%)",
                     limits = c(-60,10),
                     labels = seq(-60,10,10),
                     breaks = seq(-60,10,10)
  ) +
  xlab("Year") +
  my_ggtheme_p(leg_pos = "right",
               leg_tx_s = 8,
               leg_tl_s = 9,
               ax_tx_s = 8,
               ax_tl_s = 9,
               x_hjust = 0.5) +
  scale_color_manual("Climate change scenario",
                     values = c(wes_palette("Zissou1",type = "discrete")[1],
                                wes_palette("Zissou1",type = "discrete")[5]
                     )
  )

ggsave(
  paste0(results_path,"figures/relative_hist/all/","sq_relative_historic_cmar.png"),
  last_plot(),
  height = 3,
  width = 6
)


```

## Fig. 4. Relative to historic period (Delta)

Note that there is two options for presenting this result:

- One option is to compare each scenario to its own, in this case, aesthetics for the map should be `aes(fill = dif_2040)` and the file name `relative_historic_scen_85.png`

- A second option is to compare each scenario to the historical SQ, as William suggested. In this case, aesthetics for the map should be `aes(fill = dif_2040sq)` and the file name `relative_historic_scen_85.png`

## Figure 4 plot

```{r}

data_df %>% 
  filter(scen == "sq") %>% 
  mpa_names() %>% 
  spread(stat,value) %>% 
  # Z-score is a statistical measure that quantifies the distance between a data point and the mean of a dataset. It's expressed in terms of standard deviations.
  mutate(
    z_score = mean_esm / sd_esm,
    confidence_flag = case_when(
      abs(z_score) > 2 ~ "high",
      abs(z_score) > 1 ~ "moderate",
      TRUE ~ "low"
    ),
    pos = ifelse(ssp == "SSP1-2.6", 3.5,4.5)
  ) %>% 
  # View()
  # Check confident n
  # group_by(confidence_flag) %>%
  # mutate(n_conf = round((n()/64)*100)) %>% 
  # View()
ggplot() +
  geom_bar(
    aes(
      x = reorder(region,mean_esm),
      y = mean_esm,
      fill = ssp
    ),
    stat = "identity",
    position = "dodge"
  ) +
  geom_errorbar(
    aes(
      x = reorder(region,mean_esm),
      ymin = mean_esm - sd_esm,
      ymax = mean_esm + sd_esm,
      color = ssp
    ),
    position = "dodge",
    stat = "identity",
    linetype = "longdash",
    alpha = 0.5,
    width = 0.5
  ) +
  geom_point(
    aes(
      x = reorder(region,mean_esm),
      y = pos,
      color = ssp,
      shape = confidence_flag
    ),
    position = "dodge",
    size = 1
  ) +
  facet_wrap(~period,
             scales = "free_x"
  ) +
  scale_fill_manual("Climate change scenario",
                    values = c(wes_palette("Zissou1",type = "discrete")[1],
                               wes_palette("Zissou1",type = "discrete")[5]
                    )
                    
  ) +
  scale_color_manual("Climate change scenario",
                     values = c(wes_palette("Zissou1",type = "discrete")[1],
                                wes_palette("Zissou1",type = "discrete")[5]
                     )
  ) +
  scale_shape_discrete("Confidence level") +
  coord_flip() +
  labs(x="",
       y = "Change in biomass relative to baseline (%)")+
  my_ggtheme_p(leg_pos = "right") +
  scale_y_continuous(
    limits = c(-30,5),
    breaks = seq(-30,10,5),
    expand = c(0,0)
  ) + 
  theme(
    panel.spacing = unit(2, "lines")
  )

ggsave(
  paste0(results_path,"figures/relative_hist/all/sq_relative_historic_regional.png"),
  last_plot(),,
  height = 6,
  width = 8
)


# confidence_flag
# high	59 %
# low	17%
# moderate	23%



```

## Fig. 5A Relative to SQ (Delta Delta)

Note that this step requires the preivous one and the usage of `dif_2040sq` or `dif_2040` will impact the results.

```{r}

data_df_sq <- data_df %>% 
  filter(stat == "mean_esm") %>% 
  pivot_wider(names_from = scen, values_from = value) %>% 
  gather("scen","value",nr:rp) %>% 
  mutate(dif_per = value-sq) %>% 
  scenario_names(leng = "eng") %>% 
  filter(
    # ssp == "ssp126"
    ssp == "ssp585"
  )

map <-
  ggplot() +
  geom_sf(data = sau_sf %>% select(region = name) %>%  left_join(data_df_sq),
          aes(fill = dif_per),
          color = "grey"
  ) +
  geom_sf(data = cmar_sf %>%  left_join(data_df_sq),
          aes(fill = dif_per),
          color = "black"
  ) +
  facet_grid(period~scenario) +
  # For SQ scenario change
  scale_fill_gradient2("Differnce in biomass change rlative to status quo (%)",
                       # limits = c(-2,10),
                       # breaks = seq(-2,10,2)
                       limits = c(-2,2), # for ssp-585
                       breaks = seq(-2,2,1) # for ssp-585
  ) +
  my_ggtheme_m("Reg",
               facet_tl_s = 10,
               ax_tl_s = 8,
               ax_tx_s = 8,
               leg_tl_s = 8,
               leg_tx_s = 8) +
  theme(
    legend.key.width = unit(3, "line"),
    title = element_text(size = 4)
  )  +
  guides(fill = guide_colorbar(title.position = "top", title.hjust = 0.5))


ggsave(
  paste0(results_path,"figures/relative_scen/all/ssp585_sq_relative_map.png"),
  # paste0(results_path,"figures/relative_scen/all/ssp126_sq_relative_map.png"),
  map,
  height = 6,
  width = 10
)
```


## Fig. 5B Relative to SQ (Delta Delta)

Note that this step requires the preivous one and the usage of `dif_2040sq` or `dif_2040` will impact the results.

```{r}

data_df_sq <- data_df %>% 
  pivot_wider(names_from = scen, values_from = value) %>% 
  gather("scen","value",nr:rp) %>% 
  mutate(dif_per = value-sq
  ) %>% 
  scenario_names(leng = "eng") %>% 
  mpa_names() %>% 
  select(region:stat,scenario,dif_per) %>% 
  spread(stat,dif_per)


data_df_sq %>% 
  ggplot() +
  geom_bar(
    aes(
      x = reorder(region,desc(region)),
      y = mean_esm,
      fill = ssp
    ),
    stat = "identity",
    position = "dodge"
  ) +
  geom_errorbar(
    aes(
      x = reorder(region,desc(region)),
      ymin = mean_esm - sd_esm,
      ymax = mean_esm + sd_esm,
      color = ssp
    ),
    position = "dodge",
    stat = "identity",
    linetype = "longdash",
    alpha = 0.5,
    width = 0.5
  ) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    color = "grey"
  ) +
  # geom_point(
  #            aes(
  #              x = reorder(region,mean_esm),
  #              y = pos,
  #              color = ssp,
  #              shape = confidence_flag
  #            ),
  #            position = "dodge",
  #            size = 1
  # ) +
  facet_grid2(
    period ~ scenario,
    scales = "free_x"
  )+
  scale_fill_manual("Climate change\nscenario",
                    values = c(wes_palette("Zissou1",type = "discrete")[1],
                               wes_palette("Zissou1",type = "discrete")[5]
                    )
                    
  ) +
  scale_color_manual("Climate change\nscenario",
                     values = c(wes_palette("Zissou1",type = "discrete")[1],
                                wes_palette("Zissou1",type = "discrete")[5]
                     )
  ) +
  scale_shape_discrete("Confidence level") +
  coord_flip() +
  labs(x="",
       y = "Difference in biomass change relative to status quo (%)")+
  my_ggtheme_p(
    leg_pos = "right",
    facet_tx_s = 13,
    ax_tl_s = 14,
    ax_tx_s = 12,
    leg_tl_s = 14,
    leg_tx_s = 12
  ) +
  scale_y_continuous(
    expand = c(0,0)
  )  

ggsave(
  paste0(results_path,"figures/relative_scen/all/sq_relative_barplot.png"),
  # paste0(results_path,"figures/relative_scen/all/ssp126_sq_relative_map.png"),
  last_plot(),
  height = 7,
  width = 14
)

```


```{r}
data_df_sq %>% 
  select(-sd_esm) %>% 
  filter(
    scenario %in% c("Regulations for conservation","Regulations for fishing")
  ) %>% 
  spread(scenario,mean_esm) %>% 
  mutate(
    flag_larger = ifelse(
      `Regulations for conservation` > `Regulations for fishing`,
      "Conservation",
      "Fishing"
    ),
    # how much highwer
    dif_per = abs(`Regulations for conservation` - `Regulations for fishing`),
  ) %>% 
  View()
```

# Exploration

## Fig. Protected v.s. not protected

```{r}


gridded_data <- read_csv(paste0(data_path,"04 raw dbem results/processed/all/all_years_data_abd_esm_gridded.csv"))

grid_path <- paste0(data_path,"01 spatial/gridded/all_mpa_grid_fe.csv")

mpas_grid <- read_csv(grid_path) %>% 
  rename(mpa_abrev = mpa) %>% 
  distinct()

cmar_grid <- my_data("dbem_coords") %>% 
  filter(lat < 10 & lat > -10 & lon > -100 & lon < -70) %>% 
  left_join(mpas_grid) %>% 
  left_join(my_data("sau_index")) %>% 
  mutate(status = ifelse(is.na(status),"non_protected",status),
         region = ifelse(is.na(mpa_abrev),eez_name,mpa_abrev)
  ) %>% 
  filter(!region %in% c("Colombia (Caribbean)","Panama (Caribbean)","Costa Rica (Caribbean)","El Salvador","Nicaragua (Pacific)","Peru","Venezuela"),
         !is.na(eez_name))

gridded_df <- gridded_data %>% 
  left_join(cmar_grid) %>% 
  # group_by(year,status,region,scen,ssp,esm) %>%
  # summarise(sum = sum(value_region, na.rm = T),
  #           .groups = "drop"
  # ) %>%  
  mutate(
    time_frame = case_when(
      # year > 1994 & year < 2014 ~ "historic",
      year < 2015 ~ "historic",
      year > 2021 & year < 2040 ~ "early_2030",
      year > 2041 & year < 2060 ~ "mid_2050",
      TRUE ~ NA_character_
    )
  ) %>% 
  filter(!is.na(time_frame)) %>% 
  group_by(status,region,scen,esm,ssp,time_frame,index) %>%
  summarise(mean_time = mean(value_region, na.rm = T)) %>% 
  group_by(status,region,scen,ssp,time_frame,index) %>%
  summarise(mean_time = mean(mean_time, na.rm = T)) %>% 
  # group_by(status,region,scen,ssp,time_frame,index) %>%
  # summarise(mean_time = mean(mean_time, na.rm = T))
  # group_by(status,region,scen,ssp,time_frame) %>% 
  pivot_wider(names_from = time_frame, values_from = mean_time) %>% 
  # left_join(sq_df) %>% 
  mutate(
    dif_2030 = my_chng(historic,early_2030),
    dif_2050 = my_chng(historic,mid_2050)
  ) %>% 
  scenario_names(leng = "eng")
  # filter(!is.na(dif_2040)) %>% 
  # filter(region != "High seas") %>% 
  # summarise(
  #   across(
  #     c(dif_2030, dif_2050),
  #     list(mean_esm = ~mean(.x, na.rm = TRUE),
  #          sd_esm = ~sd(.x, na.rm = TRUE)),
  #     .names = "{.col}_{.fn}"
  #   ),
  #   .groups = "drop"
  # )# 
  # pivot_longer(cols = c("dif_2030_mean_esm": "dif_2050_sd_esm"),
  #              names_to = "period",
  #              values_to = "value") %>% 
  # mutate(
    # stat = ifelse(str_detect(period, "mean"), "mean_esm", "sd_esm"),
    # period = ifelse(str_detect(period, "2030"), "2030", "2050"),
    # ssp = ifelse(ssp == "ssp126", "SSP1-2.6","SSP5-8.5")
  # )



gridded_data_df_sq <- gridded_df %>% 
  select(status,region,scen,ssp,index,dif_2030,dif_2050) %>% 
  pivot_longer(
    cols = c("dif_2030","dif_2050"),
    names_to = "period",
    values_to = "value"
  ) %>% 
  pivot_wider(names_from = scen, values_from = value) %>% 
  gather("scen","value",nr:ri) %>% 
  mutate(dif_per = value-sq) %>% 
  scenario_names(leng = "eng")

```

```{r}

gridded_data_df_sq %>% 
  # filter(scen == "sq") %>% 
    mutate(
      eez = ifelse(str_detect(region,"Panama"),"Panama",
                         ifelse(str_detect(region,"Colombia"),"Colombia",
                                ifelse(str_detect(region,"Cuenca"),"Colombia",
                                       ifelse(str_detect(region,"Gorgona"),"Colombia",
                                ifelse(str_detect(region,"Costa"),"Costa Rica",
                                       ifelse(str_detect(region,"Ecuador"),"Ecuador","High seas"
                                       )
                                )
                         )
                                )
                         )
      )
  ) %>%
  filter(eez != "High seas") %>%
  # View()
ggplot() +
  geom_boxplot(
    aes(
      x = eez,
      y = dif_per,
      fill = scenario
    ),
    position = "dodge"
  ) +
  # facet_wrap2(
  #   ~ ssp,
  #   ncol = 1,
  #   scales = "free_x"
  # )
  facet_grid(
    ssp ~ status
  ) +
  theme_classic()+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "grey"
  )  +
  scale_fill_viridis_d("Management scenario")


ggsave(
  paste0(results_path,"figures/relative_scen/grid/sq_relative_boxplot.png"),
  last_plot(),
  height = 7,
  width = 14
)
```

### Differences between protection

```{r}

scenarios <- read_csv(paste0(data_path,"01 spatial/gridded/all_mpa_grid_fe.csv")) 

gridded_data_df_sq %>% 
  filter(status == "protected") %>% 
  left_join(scenarios,
            by ="index") %>%
  ggplot() +
  geom_point(
    aes(
      x = value,
      y = fe_prop
    ),
    position = "dodge",
    size = 1
  ) 
  
  
```


## Fig. Groups

### Data prep

```{r}
list_groups <- list.files(paste0(data_path,"04 raw dbem results/processed/groups/"),full.names = T)
list_groups <- list_groups[grep("abd", list_groups)]



scenarios_data_groups <- 
  bind_rows(
    lapply(list_groups, read_csv)
  )

group_df<- scenarios_data_groups %>% 
  select(-sd_value_esm) %>% 
  pivot_wider(names_from = period, values_from = mean_value_esm) %>% 
  mutate(
    dif_2030 = my_chng(`2014_hist`,`2030_ear`),
    dif_2050 = my_chng(`2014_hist`,`2050_mid`)
  ) %>% 
  gather("period","diff_change",`dif_2030`:`dif_2050`) %>% 
  select(scen,region,ssp,group,period,diff_change) %>% 
  pivot_wider(names_from = scen, values_from = diff_change) %>% 
  gather("scenario","value",nr:rp) %>% 
  mutate(dif_per = value-sq) %>% 
  mpa_names() %>% 
  select(region:period,scen = scenario,dif_per) %>% 
  scenario_names(leng = "eng") %>% 
  mutate(
    group = case_when(
      group == "atunes" ~ "Tuna",
      group == "dorado" ~ "Mahi-mahi",
      group == "invertebrados" ~ "Invertebrates",
      group == "jureles" ~ "Mackerels",
      group == "meros" ~ "Snappers",
      group == "others" ~ "Others",
      group == "pequeños_pelagicos" ~ "Small pelagics",
      group == "picudos" ~ "Billfishes",
      group == "tiburones" ~ "Sharks",
      TRUE ~ NA_character_
    )
  )

# unique(group_df$group)
```


```{r}

group_df %>% 
  filter(ssp =="ssp585",
         period == "dif_2050") %>% 
  # View()
  ggplot() +
  geom_bar(
    aes(
      x = reorder(region,desc(region)),
      y = dif_per,
      fill = scenario
    ),
    stat = "identity",
    position = "dodge"
  ) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    color = "grey"
  ) +
  facet_wrap2(
    ~ group,
    ncol = 3,
    scales = "free_x"
  )+
  scale_fill_viridis_d("Management scenario") +
  coord_flip() +
  labs(x="",
       y = "Difference in biomass change relative to status quo (%)")+
  my_ggtheme_p(
    leg_pos = "right",
    facet_tx_s = 13,
    ax_tl_s = 14,
    ax_tx_s = 12,
    leg_tl_s = 14,
    leg_tx_s = 12
  ) +
  scale_y_continuous(
    expand = c(0,0)
  )  + 
  theme(
    panel.spacing = unit(2, "lines")
  )


ggsave(
  # paste0(results_path,"figures/relative_scen/groups/sq_relative_barplot_126_2030.png"),
  # paste0(results_path,"figures/relative_scen/groups/sq_relative_barplot_126_2050.png"),
  # paste0(results_path,"figures/relative_scen/groups/sq_relative_barplot_585_2030.png"),
  paste0(results_path,"figures/relative_scen/groups/sq_relative_barplot_585_2050.png"),
  last_plot(),
  height =10,
  width = 12
)

```








# Old code

```{r}
#plot function
map_fx <- function(data, period_selected, ssp_selected){
  
  data_plot <- data %>% 
    filter(ssp == ssp_selected,
           period == period_selected)
  
  min_n <- round(min(data_plot$perchange,na.rm = T))-1
  max_n <- round(max(data_plot$perchange,na.rm = T))+1
  mid_n <- 0
  breaks_n <- pretty(c(min_n, max_n), n = 5)
  
  
  plot_name <- paste0(results_path,"figures/relative_hist/all/",period_selected,"_relative_historic_",ssp_selected,".png")
  # ----------- #
  
  map <-
    ggplot() +
    geom_sf(data = sau_sf %>% select(region = name) %>%  left_join(data_plot),
            aes(fill = perchange),
            color = "grey"
    ) +
    geom_sf(data = cmar_sf %>%  left_join(data_plot),
            aes(fill = perchange),
            color = "black"
    ) +
    facet_wrap(~scenario) +
    scale_fill_gradient2("Change in biomass by 2050\nrelative to baseline (%)",
                         limits = c(min_n,max_n),
                         breaks = breaks_n
    ) +
    theme(
      title = element_text(size = 4),
      legend.background = element_rect(color = "grey",linewidth = 0.5,linetype = "solid"
      )
    )  +
    my_ggtheme_m("Reg",
                 facet_tl_s = 10,
                 ax_tl_s = 10,
                 ax_tx_s = 10,
                 leg_tl_s = 2,
                 leg_tx_s = 1,
                 leg_pos = c(0.83, 0.25)) +
    guides(fill = guide_colorbar(title.position = "top", title.hjust = 0.5, title.vjust = 0.5))  
  
  
  
  ggsave(
    plot_name,
    map,
    height = 6,
    width = 8
  )
  
}

# DO NOT MODIFY
map_fx(
  data = data_df,
  ssp_selected =  "ssp585",
  period_selected = "dif_2050"
)

```

